---
title: "StockPortfolio Simulation"
author: "Bobby Fatemi"
date: "March 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stockPortfolio)
```

## Summary

This package provides functions to assist with optimizing a given portfolio. The method used is known as the fundemental theory of Portfolio Optimization (_Markowitz Mean-Variance Optimization_) and it's used to construct a portfolio that lands on the _efficient frontier_ (best portfolio given a selection of stocks). 

In the following sections, we will use \code{optimalport} to:

1. Calculate historical returns (**mu**) and risk (**sigma**) of our portfolio
2. Conduct N simulation of random walks (by day, for a year) of our portfolio using **mu** and **sigma** as inputs
3. Conduct M random samples of **mu** and **signa** and run each as new simulation

This is effectively a simulation within a simulation. What that means is we will simulate the returns of a portfolio (defined by **mu** and **sigma**), but to capture potential risk from an innacurate **mu** and **sigma** prediction, we will randomly sample those and conduct another portfolio simulation, and so on.

## Details

#### *Step 1: Get Data for Stocks*

Get historical stock prices for given stocks and calculate returns $\left(\mu_i = E(R_i)\right)$ and risk $\left(\sigma_i  = \sqrt{Variance}\right)$

#### *Step 2: Get Portfolio Weights*

Calculate the weights of each stock ($w_i$) in a optimal portfolio (efficient frontier). Note that the sum of the weights of each stock in your portfolio must equal 100%:

$$\sum_{i=1}^n w_i = w_1 + w_2 + ... + w_n = 1$$

#### *Step 3: Get Portfolio Risk and Return*

Calculate a *portfolio* risk and return pair

$$\mu_p = \sum_{i=1}^n w_i \mu_i = \sum_{i=1}^n w_i E(R_i)$$

#### *Step 4: Run Simulation (Part 1)*

Using (mu, sigma) conduct *N* random walks (1 step = 1 day) and average by day across all N random walks

#### *Step 5: Run Simulation (Part 2)*

Randomly sample another (mu, sigma) from a normal distribution and do STEP 4 again, *M times*

#### *Step 6: Derive Insights from Results*

Calculate various probabilities from our portfolio simulation

### Historical Average Risk-Returns for Each Stock

We will use the package `StockPortfolio` for a number of critical steps. Additionally, for our data wrangling and charting needs, we will use `data.table` and `googleVis`, respectively. For this, its also a good idea to turn off scientific notation for presentation of numbers.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(stockPortfolio)
library(data.table)
library(googleVis)
options(scipen = 10000)
```

##### Define Stock Tickers and Get Returns #####
Lets say we'd like a technology and banking heavy portfolio. Just because I'm lazy, let's also say we don't care about diversification and we'd like returns calculated on a daily basis going back 5 years: 

```{r,cache=TRUE}
tickers <- c("MSFT", "C", "BAC")
freq    <- "day"
start   <- Sys.Date()-365*5

stockReturns <- getReturns(tickers, freq, start = start)
```
After feeding these inputs to the function `getReturns` from the package `StockPortfolio`, we can see that we get an object of class stockModel, which is really just a list. Let's explore it:

```{r}
stockReturns
names(stockReturns)
lapply(stockReturns[["full"]],head)
```
###Weighting each Stock to get Optimal Portfolio###
The next step involves using your historical returns, calculating each stock's **mu** and **sigma**, then relying on the nobel winning Mordern Portfolio Theory introduced by Harry Markowitz to derive the best possible weighting of each stock in our portfolio. 

#####Modern Portfolio Theory#####
In a nutshell, the theory says that there is a set of reward-maximizing portfolios, based on the level of risk you take, that form what is known as *the efficient frontier*. Portfolios not on the efficient frontier can be replaced with a portfolio that either generate more reward for that level of risk, or generates same reward for lesser risk. 
> It is the set of portfolios each with the feature that no other portfolio exists with a higher expected return but with the same standard deviation of return.

We can use the mathematical fact that a portfolio's calculated return is simply a weighted average of the return of each stock, and similarly, a porfolio's standard deviation of returns is also linear with respect to the weights:

$$E(R_p) = \sum_{i}w_iE(R_i)$$ 

where R_p is the return of the portfolio, and R_i is the return on asset i. Wi is the weighting of that asset within the portfolio.

\sigma^2_p = \sum 
